Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimization Choices

### 2025-11-23 revision/response.qmd

6. The results in section 6 are very interesting. However, I was wondering about some details in 6.3: from the results shown it seems that the smoothness and squintability for the selected data examples are highly correlated, could this cause issues with the model fit shown in Table 2? It would be interesting if a similar result for CRS could be included, to see the effect of smoothness in that setting.

\textcolor{blue}{We have revised the text to include a comment on the correlation. It is moderate and does not practically affect the fit. We have also noted that it is not feasible to run CRS for the sine-wave optimization because it is not an efficient optimizer and fails to converge too often. On small dimensionality, it can be seen that it is not affected by lack of smoothness of an index. This would be expected because it is not a derivative-based optimizer, and, similarly to JSO small local fluctuations won't affect it's search.}

7. I was also wondering how the dependence of the smoothness on d can be explained, since it is not intuitive that this should change so much? (Note also that the last sentence in the caption for Table 1 seems to mix up squintability and smoothness.)

\textcolor{blue}{Table 1 is a subset of the estimated PP index, from the full table in the appendix. It happens that the three chosen ones all decrease as $n$ goes larger. This is not generally a pattern so this has been dropped from the caption.}

### 2025-11-23 jso.qmd
table 1 caption and response fixed

```{r}
#| label: tbl-smoothness-squintability
#| tbl-cap: The columns $\nu$ and $\varsigma$ show the smoothness and squintability measures for three PP indexes for the sine wave data. The splines index has the largest squint angle and is the smoothest. The skinny index has low smoothness and smallest squint angle. As dimension increases the smoothness becomes smaller but squintability doesn't change much.
dt <- tibble(shape = c(rep("pipe", 5), rep("sine", 18)), smoothness) |>
  left_join(squintability) |> 
  rename(d = n,  smooth = smoothness) |> 
  mutate(index = as.character(index),
         index = ifelse(index == "splines2d", "splines", index),
         variance = sqrt(variance)) 
dt |> 
  filter(shape == "sine", index %in% c("TIC", "splines", "skinny")) |>
  mutate(index = factor(index, levels = c("splines", "TIC", "skinny"))) |> 
  select(shape, index, d, smooth, squint) |>
  arrange(index, d) |>
  knitr::kable(
    digits = 4, format = "latex", 
    col.names = c("shape", "index", "d", "$\\nu$", "$\\varsigma$"), linesep = c("", "", "\\hline"),
    booktabs = TRUE, escape = FALSE, row.names = FALSE, align = "|llr|rr") |> 
  #column_spec(1, border_left = TRUE) |>
  column_spec(c(3,5), border_right = TRUE) 
  #row_spec(c(3, 6), extra_latex_after = "\\cline{1-5}")
  #kable_styling(font_size = 10)   
```

>>>>

```{r}
#| label: tbl-smoothness-squintability
#| tbl-cap: The columns $\nu$ and $\varsigma$ show the smoothness and squintability measures for PP indexes for the sine wave data. Only three are shown for illustration, and a full table can be found in Supplementary materials. The splines index has the largest squint angle and is the smoothest. The skinny index has low smoothness and smallest squint angle. 
dt <- tibble(shape = c(rep("pipe", 5), rep("sine", 18)), smoothness) |>
  left_join(squintability) |> 
  rename(d = n,  smooth = smoothness) |> 
  mutate(index = as.character(index),
         index = ifelse(index == "splines2d", "splines", index),
         variance = sqrt(variance)) 
dt |> 
  filter(shape == "sine", index %in% c("TIC", "splines", "skinny")) |>
  mutate(index = factor(index, levels = c("splines", "TIC", "skinny"))) |> 
  select(shape, index, d, smooth, squint) |>
  arrange(index, d) |>
  knitr::kable(
    digits = 4, format = "latex", 
    col.names = c("shape", "index", "d", "$\\nu$", "$\\varsigma$"), linesep = c("", "", "\\hline"),
    booktabs = TRUE, escape = FALSE, row.names = FALSE, align = "|llr|rr") |> 
  #column_spec(1, border_left = TRUE) |>
  column_spec(c(3,5), border_right = TRUE) 
  #row_spec(c(3, 6), extra_latex_after = "\\cline{1-5}")
  #kable_styling(font_size = 10)   
```

### 2025-11-23 jso.tex
table 1 caption and response fixed

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colframe=quarto-callout-note-color-frame, toprule=.15mm, arc=.35mm, opacityback=0, opacitybacktitle=0.6, bottomrule=.15mm, breakable, left=2mm, coltitle=black, bottomtitle=1mm, toptitle=1mm, colbacktitle=quarto-callout-note-color!10!white, leftrule=.75mm, titlerule=0mm]

@@

\caption{\label{tbl-smoothness-squintability}The columns \(\nu\) and
\(\varsigma\) show the smoothness and squintability measures for three
PP indexes for the sine wave data. The splines index has the largest
squint angle and is the smoothest. The skinny index has low smoothness
and smallest squint angle. As dimension increases the smoothness becomes
smaller but squintability doesn't change much.}

>>>>

\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, rightrule=.15mm, colback=white, toptitle=1mm, opacityback=0, arc=.35mm, left=2mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, bottomrule=.15mm, coltitle=black, bottomtitle=1mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, leftrule=.75mm, toprule=.15mm, breakable, colbacktitle=quarto-callout-note-color!10!white]

@@

\caption{\label{tbl-smoothness-squintability}The columns \(\nu\) and
\(\varsigma\) show the smoothness and squintability measures for PP
indexes for the sine wave data. Only three are shown for illustration,
and a full table can be found in Supplementary materials. The splines
index has the largest squint angle and is the smoothest. The skinny
index has low smoothness and smallest squint angle.}
